---
title: "20240530_SDR005_GO_analysis-V2"
output: html_document
date: "2024-05-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#Load libraries


```{r, fig.width = 6, fig.height = 10}

library(Seurat)
library(dplyr)
library(ggplot2)
library(Matrix)
library(reshape2)
library(data.table)
library(stringr)
library(tidyr)
library(scales)
library(rtracklayer)
library(data.table)
library(readr)
library(pheatmap)
library(RColorBrewer)
library(topGO)
library(org.Hs.eg.db)
library(purrr)



file_save_dir <- "/Users/dominik.lindenhofer/ownCloud/Shared_project_folder/Paper_drafts/2022_SDR_Paper/Source_data/Individual_files_all"


```




#load data 



#Load in variants for which to filter for from Chi-square testing LZ -DZ

```{r, fig.width = 6, fig.height = 10}

cDNA_data_Seurat <- readRDS("SDR005_Combined_Geno_full_clustered.rds")


Chi_FL7_one <- read.csv("SDR005_Stat_testing/Chi_FL7_Cell_Types_all.csv") %>% dplyr::select(-X) %>% mutate(sample = "FL7")
Chi_FL11_one <- read.csv("SDR005_Stat_testing/Chi_FL11_Cell_Types_all.csv") %>% dplyr::select(-X) %>% mutate(sample = "FL11")
Chi_GCB4_one <- read.csv("SDR005_Stat_testing/Chi_GCB4_Cell_Types_all.csv")  %>% dplyr::select(-X) %>% mutate(sample = "GCB4")


Chi_comb_one <- rbind(Chi_FL7_one, Chi_FL11_one, Chi_GCB4_one)


#Remove WT - WT 1 ones - filter for DZ and LZ 

Chi_comb_one <- Chi_comb_one %>% 
  filter(State1 == "DZ" & State2 == "LZ") %>%
  filter(!(State1_0_perc == 1 & State2_0_perc == 1)) %>%
  mutate(across(everything(), ~ifelse(is.na(.), 0, .)))

#set_cut_off for number of to filter out low variants 


#Add info
Chi_comb_one <- Chi_comb_one  %>% 
  mutate(State1_edit_count = State1_1_count + State1_2_count,
         State2_edit_count = State2_1_count + State2_2_count,
         State1_edit_perc = State1_1_perc + State1_2_perc,
         State2_edit_perc = State2_1_perc + State2_2_perc,) %>%
  mutate(edit_count_tot = State1_edit_count + State2_edit_count) %>%
  mutate(Significant = if_else(P_val_adj < 0.05, TRUE, FALSE))

#Apply cutoff for variants

cut_off = 30
Chi_comb_one <- Chi_comb_one  %>% filter(edit_count_tot > cut_off)

cut_off = 0.05
Chi_comb_one <- Chi_comb_one  %>% filter(State1_edit_perc > cut_off |
                                         State2_edit_perc > cut_off )


#Add delta Chi

Chi_comb_one <- Chi_comb_one %>% mutate(delta_0_perc = State1_0_perc - State2_0_perc,
                                        delta_1_perc = State1_1_perc - State2_1_perc,
                                        delta_2_perc = State1_2_perc - State2_2_perc)



#Add Gene ID

Geno_meta <- cDNA_data_Seurat@assays$Geno_filtered@meta.data
to_add <- Geno_meta %>% dplyr::select(ID_Geno, Gene_annot, Consequence)
Chi_comb_one <- Chi_comb_one %>% left_join(to_add, by = c("Variant" = "ID_Geno" ))

#Add to highlight

Chi_comb_one  <- Chi_comb_one %>% mutate(target_to_plot_0 = case_when(Significant == TRUE & delta_0_perc > 0 ~ State1,
                                                                      Significant == TRUE & delta_0_perc < 0 ~ State2))


Chi_comb_one  <- Chi_comb_one %>% mutate(target_to_plot_1 = case_when(Significant == TRUE & delta_1_perc > 0 ~ State1,
                                                                      Significant == TRUE & delta_1_perc < 0 ~ State2))


Chi_comb_one  <- Chi_comb_one %>% mutate(target_to_plot_2 = case_when(Significant == TRUE & delta_2_perc > 0 ~ State1,
                                                                      Significant == TRUE & delta_2_perc < 0 ~ State2))


to_order <- c("FL7", "FL11", "GCB4")

Chi_comb_one$sample <- factor(Chi_comb_one$sample, levels = to_order)


#Exclude IG variants

Chi_comb_one_wo_IG <- Chi_comb_one %>% mutate(Gene_annot = if_else(str_detect(Gene_annot, "^IG"), NA_character_, Gene_annot)) %>%   filter(!is.na(Gene_annot))


#Select significant
#All
Chi_comb_one_sel <- Chi_comb_one %>% filter(Significant == TRUE)

#no IG included
Chi_comb_one_wo_IG_sel <- Chi_comb_one_wo_IG %>% filter(Significant == TRUE)


```




#Include IG variants in this analysis 

#Laod DE output - all variants
#Subset for only CHI-DE ones from above


```{r, fig.width = 6, fig.height = 10}


#_________________________________________________________________________

#Load data
#cDNA_data_Seurat <- readRDS("SDR005_Combined_Geno_full_clustered.rds")

metadata <- cDNA_data_Seurat@meta.data


#Get all matrices
RNA_mat <- cDNA_data_Seurat@assays$RNA$data %>% t() %>% as.data.frame() %>% exp() -1
GT_mat <- cDNA_data_Seurat@assays$Geno_filtered$counts %>% as.data.frame()
Geno_meta <- cDNA_data_Seurat@assays$Geno_filtered@meta.data

Geno_list <- cDNA_data_Seurat@assays$Geno_filtered@misc
list2env(Geno_list, envir = .GlobalEnv)



#Load - all variants 

DE_test_comb <- read.csv("SDR005_Stat_testing_MAST/SDR005_DE_test_out_DZ_LZ_comb.csv")

#Fix gene names
DE_test_comb <- DE_test_comb %>% mutate(Gene = gsub("IGLC1-IGLL5", "IGLL5", Gene))


#Subset for variants in Chi_comb_one_sel

to_filter <- unique(Chi_comb_one_sel$Variant)

DE_test_comb <- DE_test_comb %>% filter(ID_Geno %in% to_filter)


#Select significant ones
DE_test_comb_sel <- DE_test_comb %>% filter(p_val_adj < 0.05,  
                                                        pct.1 > 0.25 & pct.2 > 0.25)
#Check most frequent effects
sum <- DE_test_comb_sel  %>% group_by(Sample, State, Gene) %>% summarise(count = n())
sum2 <- DE_test_comb_sel %>% group_by(Sample, Gene) %>% summarise(count = n())
sum3 <- DE_test_comb_sel %>% group_by(Gene_annot, Gene) %>% summarise(count = n())
sum4 <- DE_test_comb_sel %>% group_by(Sample, ID_Geno) %>% summarise(count = n())
sum5 <- DE_test_comb_sel %>% group_by(Sample, State, Gene, Gene_annot) %>% summarise(count = n())



#filter out IG genes
DE_test_comb_wo_IG <- DE_test_comb %>% mutate(Gene_annot = if_else(str_detect(Gene_annot, "^IG"), NA_character_, Gene_annot)) %>%   filter(!is.na(Gene_annot))
  
  

DE_test_comb_wo_IG_sel <- DE_test_comb_wo_IG %>% filter(p_val_adj < 0.05,  
                                                        pct.1 > 0.25 & pct.2 > 0.25)



  
```




#Make summarized count plots - all variants that are Chi-DE between samples - cut 50



```{r, fig.width = 10, fig.height = 6}

#to_filter <- c("BCL2", "HIST1H1D", "DTX1", "HIST1HE", "BCL6", "TNFRSF14")
group_colors_samples <- c("FL7" = "#FC6736", FL11 = "#FFB0B0", GCB4 = "#0C2D57")

to_filter_count <- DE_test_comb_sel %>% group_by(Sample, Gene) %>% summarise(count = n())



to_filter_count <- DE_test_comb_sel %>% group_by(Gene) %>% summarise(count = n()) %>% filter(count > 26)

sum5 <- DE_test_comb_sel %>% group_by(Sample, Gene) %>% summarise(count = n()) %>% filter(Gene %in% to_filter_count$Gene)

sum5$Gene

# Filter the sum5 dataframe to include only the genes of interest
sum5_sel <- sum5 
write.csv(sum5_sel, "Genes_to_filter.csv")


#sum5_sel <- sum5

# Define the desired order of the samples
to_order_sample <- c("FL7", "FL11", "GCB4")

# Summarize the count of each gene and arrange in descending order
to_order_gene <- sum5_sel %>%
  group_by(Gene) %>%
  summarize(count = sum(count)) %>%
  arrange(desc(count))

# Convert the Gene column to a factor to control the order
to_order_gene$Gene <- as.factor(to_order_gene$Gene)

# Reorder the Sample column in sum5_sel according to to_order_sample
sum5_sel$Sample <- factor(sum5_sel$Sample, levels = to_order_sample)

# Reorder the Gene column in sum5_sel according to to_order_gene
sum5_sel$Gene <- factor(sum5_sel$Gene, levels = to_order_gene$Gene)

  
title = "Gene_effect_count_all_variants.pdf"
p <- sum5_sel %>% 
    ggplot(aes(x = Gene, y = count, fill = Sample)) + 
      geom_bar(position = "stack", stat = "identity") + 
    	theme_classic() +
  	theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  #theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)+
        scale_fill_manual(values = group_colors_samples)+
    guides(fill = FALSE)

p

ggsave(filename = title, path = "SDR005_Stat_testing_GO", plot = p, device = "pdf", scale = 1,
       width = 30, height = 15, units = "cm", dpi = 300)



to_file <- sum5_sel 
to_file <- to_file %>% mutate(Sample = case_when(Sample == "FL7" ~ "FL1",
                                                 Sample == "FL11" ~ "FL2",
                                                 Sample == "GCB4" ~ "GCB1"))


file_name = "Fig4_S9_e_Gene_effect_count_all_variants.csv"
write.csv(to_file, paste0(file_save_dir, "/", file_name), row.names = FALSE) 




```






#Make dotplot of expression - top DE genes 


```{r}

#Check RNA_mat
rowSums(RNA_mat)

to_add <- metadata %>% dplyr::select(cell_merge, Tumor_ID, CellTypes)

RNA_mat_mod <- RNA_mat %>% mutate(cell = rownames(RNA_mat))
                                                  
RNA_mat_mod_res <- RNA_mat_mod %>% left_join(to_add, by = c("cell" = "cell_merge")) %>% reshape2::melt()

colnames(RNA_mat_mod_res) <- c("cell", "Tumor_ID", "State", "ID", "nUMIs")

nCells_Exp <- metadata %>% group_by(Tumor_ID, CellTypes) %>% summarise(nCells = n_distinct(cell_merge))

#Make average expression and fraction of cells detected per gene - wo normalization 

RNA_mat_mod_res_expr_all <- RNA_mat_mod_res %>% 
  dplyr::group_by(ID, Tumor_ID, State) %>% 
  dplyr::summarize(avg_nUMIs = mean(nUMIs),
                   n_cells = sum(nUMIs > 0), 
                   .groups = "drop") %>% 
    dplyr::left_join(nCells_Exp, by = c("Tumor_ID", "State" = "CellTypes" )) %>%
dplyr::mutate(avg_nUMIs_log1p = log1p(avg_nUMIs),
                frac_cells = n_cells/nCells)


#Set filter for genes - from barplot above

to_filter_count <- DE_test_comb_sel %>% group_by(Gene) %>% summarise(count = n()) %>% filter(count > 30)

sum5 <- DE_test_comb_sel %>% group_by(Sample, Gene) %>% summarise(count = n()) %>% filter(Gene %in% to_filter_count$Gene)
sum5_sel <- sum5 
genes_to_filter <- unique(sum5_sel$Gene)
length(genes_to_filter)



RNA_mat_mod_res_expr_all_filter <- RNA_mat_mod_res_expr_all %>% filter(ID %in% genes_to_filter &
                                                                       State %in% c("DZ", "LZ"))

length(unique(RNA_mat_mod_res_expr_all_filter$ID))
#RNA_mat_mod_res_expr_all_filter$ID <- factor(RNA_mat_mod_res_expr_all_filter$ID, levels = levels(genes_to_filter))



#Make unique column for plotting

RNA_mat_mod_res_expr_all_filter <- RNA_mat_mod_res_expr_all_filter %>% mutate(sample_State = str_c(Tumor_ID, State, sep = "_"))

#Set order for plotting

to_order <- c("FL7_DZ", "FL7_LZ", "FL11_DZ", "FL11_LZ", "GCB4_DZ", "GCB4_LZ")

RNA_mat_mod_res_expr_all_filter$sample_State <- factor(RNA_mat_mod_res_expr_all_filter$sample_State, levels =  to_order )

max_plot <- max(RNA_mat_mod_res_expr_all_filter$avg_nUMIs_log1p)



pl1 <-  RNA_mat_mod_res_expr_all_filter  %>%
  ggplot(aes(y = ID,  x = factor(sample_State))) + 
  geom_tile(aes(fill = avg_nUMIs_log1p)) +  
  scale_fill_continuous(low = "white", high = "white")

pl1



pl2<- pl1  +   
  geom_point(aes(color=avg_nUMIs_log1p, size = frac_cells)) +    
  scale_colour_gradient(low = "grey", high = "#990000", limits = c(0,max_plot)) +  
  #scale_colour_gradient2(low = "#990000", mid = "grey", high = "#336600", midpoint = 0, limits = c(0,3))+     
  scale_size(breaks = c(0, 0.25, 0.5, 0.75, 1), range = c(0, 4), limits= c(0,1))+ 
  # scale_size(breaks = c(0.5, 0.75, 1), range = c(0, 10), limits= c(0,1))+ 
  theme_bw()

pl2


ggsave(filename = "Dotplot_top_DE_genes_for_figure.pdf", path = "SDR005_Stat_testing_GO", plot = pl2, device = "pdf", scale = 1,width = 15, height = 40, units = "cm", dpi = 300)

```




#Make heatmap of the plot above


```{r}



to_plot_FL7_DZ <- RNA_mat_mod_res_expr_all_filter %>% filter(sample_State == "FL7_DZ")
to_plot_FL7_LZ <- RNA_mat_mod_res_expr_all_filter %>% filter(sample_State == "FL7_LZ")
to_plot_FL11_DZ <- RNA_mat_mod_res_expr_all_filter %>% filter(sample_State == "FL11_DZ")
to_plot_FL11_LZ <- RNA_mat_mod_res_expr_all_filter %>% filter(sample_State == "FL11_LZ")
to_plot_GCB4_DZ <- RNA_mat_mod_res_expr_all_filter %>% filter(sample_State == "GCB4_DZ")
to_plot_GCB4_LZ <- RNA_mat_mod_res_expr_all_filter %>% filter(sample_State == "GCB4_LZ")


to_plot <- data.frame(FL7_DZ = to_plot_FL7_DZ$avg_nUMIs, FL7_LZ = to_plot_FL7_LZ$avg_nUMIs,
                      FL11_DZ = to_plot_FL11_DZ$avg_nUMIs, FL11_LZ = to_plot_FL11_LZ$avg_nUMIs,
                      GCB4_DZ = to_plot_GCB4_DZ$avg_nUMIs, GCB4_LZ = to_plot_GCB4_LZ$avg_nUMIs)



rownames(to_plot) <- to_plot_FL7_DZ$ID





#_________________________________________________________________________

#Plot heatmap

#Set Breaks for dataset
myBreaks = seq(-2, 2, by = 0.1)
#breaksList <- seq(min(RNA_seq_iPSCs), max(RNA_seq_iPSCs), length.out = 50)

#Set colors for dataset
#myColor <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(length(breaksList))
myColor = colorRampPalette(c("#F6F5F2", "yellow3", "#337357"))(length(myBreaks))


title = "Heatmap_top_DE_genes_for_figure.pdf"
p <- pheatmap(to_plot,
         #color = inferno(255),
         #color = brewer.pal(11, "RdBu"),
        color = myColor,
         scale = "row",
         #scale = "none",
         cluster_cols = FALSE,
         cluster_rows = TRUE,
         border_color = NA,
         show_colnames = TRUE,
         show_rownames = TRUE,
         drop_levels = TRUE,
         #annotation_row = mat_row,
         #annotation_col = mat_col,
         #annotation_colors = group_colors_heatmap ,
         #cutree_cols = 3,
         #cutree_rows = 2
         breaks = myBreaks,
         main = title
)


ggsave(filename = title, path = "SDR005_Stat_testing_GO", plot = p, device = "pdf", scale = 1,
       width = 15, height = 15, units = "cm", dpi = 300)

p




to_file <- to_plot %>% mutate(Gene= rownames(to_plot))

file_name = "Fig4_M_g_Heatmap_top_DE_genes_for_figure.csv"
write.csv(to_file, paste0(file_save_dir, "/", file_name), row.names = FALSE) 



```



#Make heatmap with all genes LZ/DZ to check general expression of genes in those states


```{r}


#Prepare RNA_mat and add cell column for filtering 

rowSums(RNA_mat)

to_add <- metadata %>% dplyr::select(cell_merge, Tumor_ID, CellTypes)

RNA_mat_mod <- RNA_mat %>% mutate(cell = rownames(RNA_mat))
                                                  
RNA_mat_mod_res <- RNA_mat_mod %>% left_join(to_add, by = c("cell" = "cell_merge")) %>% reshape2::melt()

colnames(RNA_mat_mod_res) <- c("cell", "Tumor_ID", "State", "ID", "nUMIs")

nCells_Exp <- metadata %>% group_by(Tumor_ID, CellTypes) %>% summarise(nCells = n_distinct(cell_merge))

#Make average expression and fraction of cells detected per gene - wo normalization 

RNA_mat_mod_res_expr_all <- RNA_mat_mod_res %>% 
  dplyr::group_by(ID, Tumor_ID, State) %>% 
  dplyr::summarize(avg_nUMIs = mean(nUMIs),
                   n_cells = sum(nUMIs > 0), 
                   .groups = "drop") %>% 
    dplyr::left_join(nCells_Exp, by = c("Tumor_ID", "State" = "CellTypes" )) %>%
dplyr::mutate(avg_nUMIs_log1p = log1p(avg_nUMIs),
                frac_cells = n_cells/nCells)


RNA_mat_mod_res_expr_all_filter <- RNA_mat_mod_res_expr_all %>% filter(State %in% c("DZ", "LZ"))
                                                                       





#Make unique column for plotting

RNA_mat_mod_res_expr_all_filter <- RNA_mat_mod_res_expr_all_filter %>% mutate(sample_State = str_c(Tumor_ID, State, sep = "_"))

#Set order for plotting

to_order <- c("FL7_DZ", "FL7_LZ", "FL11_DZ", "FL11_LZ", "GCB4_DZ", "GCB4_LZ")

RNA_mat_mod_res_expr_all_filter$sample_State <- factor(RNA_mat_mod_res_expr_all_filter$sample_State, levels =  to_order )

max_plot <- max(RNA_mat_mod_res_expr_all_filter$avg_nUMIs_log1p)



RNA_mat_mod_res_expr_all_filter <- RNA_mat_mod_res_expr_all_filter %>% dplyr::select(sample_State, ID, avg_nUMIs)

#Reshape


RNA_mat_mod_res_expr_all_filter_wide <- RNA_mat_mod_res_expr_all_filter %>% pivot_wider(names_from = ID, values_from = avg_nUMIs) %>% ungroup() %>% as.data.frame()

rownames(RNA_mat_mod_res_expr_all_filter_wide) <- RNA_mat_mod_res_expr_all_filter_wide$sample_State


RNA_mat_mod_res_expr_all_filter_wide <- RNA_mat_mod_res_expr_all_filter_wide  %>% dplyr::select(-sample_State)

#Remove columns with all 0
to_remove <- colSums(RNA_mat_mod_res_expr_all_filter_wide) %>% as.data.frame()
to_remove <- to_remove %>% filter(. != 0)

RNA_mat_mod_res_expr_all_filter_wide <- RNA_mat_mod_res_expr_all_filter_wide[, rownames(to_remove)]




to_plot <- RNA_mat_mod_res_expr_all_filter_wide

#_________________________________________________________________________

#Plot heatmap

#Set Breaks for dataset
myBreaks = seq(-2, 2, by = 0.1)
#breaksList <- seq(min(RNA_seq_iPSCs), max(RNA_seq_iPSCs), length.out = 50)

#Set colors for dataset
#myColor <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(length(breaksList))
myColor = colorRampPalette(c("#F6F5F2", "yellow3", "#337357"))(length(myBreaks))


title = "Heatmap_all_genes_LZ_DZ.pdf"
p <- pheatmap(to_plot,
         #color = inferno(255),
         #color = brewer.pal(11, "RdBu"),
        color = myColor,
         scale = "column",
         #scale = "none",
         #cluster_cols = TRUE,
         cluster_rows = FALSE,
         border_color = NA,
         show_colnames = TRUE,
         show_rownames = TRUE,
         drop_levels = TRUE,
         #annotation_row = mat_row,
         #annotation_col = mat_col,
         #annotation_colors = group_colors_heatmap ,
         cutree_cols = 5,
         #cutree_rows = 2
         breaks = myBreaks,
         main = title
)


ggsave(filename = title, path = "SDR005_Stat_testing_GO", plot = p, device = "pdf", scale = 1,
       width = 100, height = 15, units = "cm", dpi = 300)

p



```



#Make GO term analysis of TOP DE genes in all samples


```{r}


data(geneList)
sum(topDiffGenes(geneList))

#All combined - only top ones 

all.genes <- unique(DE_test_comb$Gene)

#Subset DE genes - take from subset above 

#to_filter <- c("BCL2", "HIST1H1D", "DTX1", "HIST1HE", "BCL6", "TNFRSF14")
to_filter_count <- DE_test_comb_sel %>% group_by(Gene) %>% summarise(count = n())



#Define missing ones 

missing.genes <- setdiff(all.genes, to_filter_count$Gene)

df_missing.genes <- data.frame(Gene = missing.genes) %>% mutate(count = 0)


combined_df <- rbind(to_filter_count, df_missing.genes)

p_values <- combined_df %>% dplyr::select(count) %>% pull(1)
genes <- combined_df %>% dplyr::select(Gene) %>% pull(1)
names(p_values) <- genes


# Function to map gene symbols to Entrez IDs
getEntrezIDs <- function(symbols) {
  suppressWarnings(AnnotationDbi::select(org.Hs.eg.db, keys = symbols, columns = "ENTREZID", keytype = "SYMBOL"))
}

#Get EntrezID numbers
mappedGenes <- getEntrezIDs(genes)
mappedGenes <- mappedGenes[!is.na(mappedGenes$ENTREZID),]
mappedGenes$ENTREZID <- as.character(mappedGenes$ENTREZID)  

#Prepare GO data
#genes_list <- mappedGenes$ENTREZID
#geneID2GO <- annFUN.org(whichOnto = "BP", feasibleGenes = mappedGenes$ENTREZID, mapping = "org.Hs.eg.db", ID = "entrez")  # for Biological Process
#geneID2GO <- annFUN.gene2GO(whichOnto = "BP", feasibleGenes = mappedGenes$ENTREZID, gene2GO = geneID2GO)  # for Biological Process


# Filtering function to select significant genes
geneSelectionFun <- function(pvalue) {
  return(pvalue > 26)
}

#BP
GOdata <- new("topGOdata",
              ontology = "BP",
              allGenes = p_values,
              geneSel = geneSelectionFun,
              nodeSize = 10,
              annot = annFUN.org, 
              mapping = "org.Hs.eg.db",
              ID = "symbol") 


genes(GOdata)
numGenes(GOdata)
sigGenes(GOdata)
numSigGenes(GOdata)

#Run tests

# Running the enrichment test using the 'weight01' algorithm with Fisher's exact test
resultWeight01Fisher <- runTest(GOdata, algorithm = "weight01", statistic = "fisher")
resultClassicFisher <- runTest(GOdata, algorithm = "classic", statistic = "fisher")
resultWeight01KS <- runTest(GOdata, algorithm = "weight01", statistic = "ks")
resultClassicKS <- runTest(GOdata, algorithm = "classic", statistic = "ks")



# Summarizing the results
resultsTable <- GenTable(GOdata,
                         Weight01Fisher = resultWeight01Fisher,
                         ClassicFisher = resultClassicFisher,
                         Weight01KS = resultWeight01KS,
                         ClassicKS = resultClassicKS, 
                         orderBy = "Weight01Fisher",  # Order results by the weight01 Fisher test
                         ranksOf = "ClassicFisher",
                         topNodes = 30)  # Get the top 10 significant nodes/terms


pvalWeight <- score(resultWeight01Fisher)
hist(pvalWeight, 50, xlab = "p-values")
pvalClassic <- score(resultClassicFisher)
hist(pvalClassic, 50, xlab = "p-values")
cor(pvalWeight, pvalClassic)
geneData(resultWeight01Fisher)

genes_in_term <- genesInTerm(GOdata)[["GO:0050853"]]

#Plot GO term

resultsTable$Weight01Fisher <- as.numeric(resultsTable$Weight01Fisher)

resultsTable <- resultsTable[c(1:10),]

resultsTable$Term <- factor(resultsTable$Term, levels = resultsTable$Term )



title = "GO_BP_filtered_variants.pdf"
p <- resultsTable %>% 
    ggplot(aes(x = Term, y = -log10(Weight01Fisher))) + 
      geom_bar(stat = "identity") + 
    	theme_classic() +
  	theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  #theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)+
        #scale_fill_manual(values = group_colors_samples)+
    guides(fill = FALSE)

p

ggsave(filename = title, path = "SDR005_Stat_testing_GO", plot = p, device = "pdf", scale = 1,
       width = 15, height = 15, units = "cm", dpi = 300)



to_file <- resultsTable %>% dplyr::select(GO.ID, Term, P_val = Weight01Fisher)

file_name = "Fig4_M_h_GO_BP_filtered_variants.csv"
write.csv(to_file, paste0(file_save_dir, "/", file_name), row.names = FALSE) 



```



#Subset cells with more and less variants per cell - nCells edit 


```{r}

plot_save_dir <- "/Users/dominik.lindenhofer/Documents/Data_Local/Shared_Donnacha_SDR005!/SDR005_data_share/20240205_SDR005_SDRranger_gDNA-V4_cDNA-V5/SDR005_Stat_testing_GO/01_HIGH_LOW"


#Load data
cDNA_data_Seurat_DZ_LZ_split <- readRDS("SDR005_Combined_Geno_full_clustered_Gclustered_DZ_LZ_sub_split.rds")

#__________________________

#Make loop

i = "FL7"
x = "ACTB"

Gene = "BLK"

#Genes <- sum5_sel$Gene

Genes <- genes_in_term 

count_GT_list <- list()
to_plot_list_list <- list()

for (i in names(cDNA_data_Seurat_DZ_LZ_split)) {
metadata_i <- cDNA_data_Seurat_DZ_LZ_split[[i]]@meta.data

GT_mat_i <- cDNA_data_Seurat_DZ_LZ_split[[i]]@assays$Geno_filtered$counts %>% as.data.frame()
RNA_mat_i <- cDNA_data_Seurat_DZ_LZ_split[[i]]@assays$RNA$data %>% t() %>% as.data.frame() %>% exp() -1


#cells DZ and LZ

cells_DZ <- metadata_i %>% filter(CellTypes == "DZ") %>% pull(cell_merge)
cells_LZ <- metadata_i %>% filter(CellTypes == "LZ") %>% pull(cell_merge)


#Count variants per cell
count_GT_i <- GT_mat_i  %>%
  apply(2, function(x) table(factor(x, levels = c("0", "1", "2")))) %>%
  t() %>%
  as.data.frame()

colnames(count_GT_i) <- c("REF", "HET", "ALT")
count_GT_i <- count_GT_i %>% mutate(HET_ALT = HET + ALT,
                                    cell_merge = rownames(count_GT_i)) 

#Select gene for visualization 
RNA_mat_i_sel <- RNA_mat_i[,Genes]
RNA_mat_i_sel <- RNA_mat_i %>% mutate(cell_merge = rownames(RNA_mat_i))

to_add_i <- metadata_i %>% dplyr::select(CellTypes, cell_merge)

#Merge with Gene info
count_GT_i <- count_GT_i %>% left_join(RNA_mat_i_sel, by = "cell_merge") %>% 
                             left_join(to_add_i, by = "cell_merge") %>% 
                             mutate(sample = i)



#Assign top and bottom variant cells

top_thr_i_DZ <- count_GT_i %>% filter(CellTypes == "DZ") %>% pull(HET_ALT) %>% quantile(0.8)
bot_thr_i_DZ <- count_GT_i %>% filter(CellTypes == "DZ") %>% pull(HET_ALT) %>% quantile(0.2)                                      

top_thr_i_LZ <- count_GT_i %>% filter(CellTypes == "LZ") %>% pull(HET_ALT) %>% quantile(0.8)
bot_thr_i_LZ <- count_GT_i %>% filter(CellTypes == "LZ") %>% pull(HET_ALT) %>% quantile(0.2)                                       


count_GT_i <- count_GT_i %>% mutate(Class = case_when(CellTypes == "DZ" & HET_ALT > top_thr_i_DZ ~ "HIGH",
                                                      CellTypes == "DZ" & HET_ALT < bot_thr_i_DZ ~ "LOW",
                                                      CellTypes == "LZ" & HET_ALT > top_thr_i_LZ ~ "HIGH",
                                                      CellTypes == "LZ" & HET_ALT < bot_thr_i_LZ ~ "LOW"))


count_GT_i <- count_GT_i %>% mutate(ID = str_c(sample, CellTypes, Class, sep = "_"))

#Calculate avg expression for each gene in top bottom

to_plot_list <- list()
for(x in Genes){
  
count_GT_i_x <- count_GT_i %>%
    filter(Class == "HIGH" | Class == "LOW") %>%
    dplyr::select(ID, !!sym(x)) %>%
    group_by(ID) %>%
    summarise(avg = mean(.data[[x]], na.rm = TRUE)) %>% # Using .data to access variable x dynamically
    mutate(sample = i,
           Gene = x)

to_plot_list[[x]] <- count_GT_i_x

}

to_plot_list_list[[i]] <- to_plot_list

count_GT_list[[i]] <- count_GT_i

}


#Plot heatmap all genes
flattened_to_plot <- map(to_plot_list_list, bind_rows)
to_plot <- bind_rows(flattened_to_plot, .id = "sample")

to_plot_wide <- to_plot %>%
  pivot_wider(names_from = Gene, values_from = avg)

# Set 'ID' as row names and remove 'ID' and 'sample' columns
to_rownames <- to_plot_wide$ID  # Extract the ID column for row names
to_plot_wide <- to_plot_wide %>%
  dplyr::select(-ID, -sample) %>% 
  as.data.frame()  # Convert to data frame for setting row names
rownames(to_plot_wide) <- to_rownames  # Assign row names


myBreaks = seq(-2, 2, by = 0.1)
#breaksList <- seq(min(RNA_seq_iPSCs), max(RNA_seq_iPSCs), length.out = 50)

#Set colors for dataset
#myColor <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(length(breaksList))
myColor = colorRampPalette(c("#F6F5F2", "yellow3", "#337357"))(length(myBreaks))





title = "SDR005_HIGH_LOW_heatmap_LZ_DZ.pdf"
p <- pheatmap(to_plot_wide,
         #color = inferno(255),
         #color = brewer.pal(11, "RdBu"),
        color = myColor,
         scale = "column",
         #scale = "none",
         cluster_cols = TRUE,
         cluster_rows = FALSE,
         border_color = NA,
         show_colnames = TRUE,
         show_rownames = TRUE,
         drop_levels = TRUE,
         #annotation_row = mat_row,
         #annotation_col = mat_col,
         #annotation_colors = group_colors_heatmap ,
         #cutree_cols = 3,
         #cutree_rows = 2
         #breaks = myBreaks,
         main = title
)


ggsave(filename = title, path = "SDR005_Stat_testing_GO", plot = p, device = "pdf", scale = 1,
       width = 15, height = 15, units = "cm", dpi = 300)

p





to_file <- to_plot_wide %>% mutate(ID = rownames(to_plot_wide ))


file_name = "Fig4_M_i_SDR005_HIGH_LOW_heatmap_LZ_DZs.csv"
write.csv(to_file, paste0(file_save_dir, "/", file_name), row.names = FALSE) 







```


