---
title: "20230328_SDR002_SDRranger"
author: "Dominik Lindenhofer"
date: "2024-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages


```{r}

library(Seurat)
library(dplyr)
library(ggplot2)
library(Matrix)
library(reshape2)
library(data.table)
library(stringr)
library(tidyr)
library(scales)
library(rtracklayer)
library(here)


plot_save_dir <- here("SDR002_240_240_Plots")
path_dir_gDNA_reads <- here("SDRranger_gDNA_240/raw_reads_bc_matrix")
path_dir_cDNA_reads <- here("SDRranger_cDNA_240/raw_reads_bc_matrix")
path_dir_cDNA_umis <- here("SDRranger_cDNA_240/raw_umis_bc_matrix")
panel_size = 240

```

# Use non filtered data as input

#Construct gDNA metadata table - from gDNA reads

```{r}

#load read data sets 

gDNA_data_features_reads <- read.delim(paste0(path_dir_gDNA_reads, "/features.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
gDNA_data_features_reads <- sub("_chrom", "\\1", gDNA_data_features_reads)
gDNA_data_barcodes_reads <- read.delim(paste0(path_dir_gDNA_reads,"/barcodes.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
gDNA_mat_reads <- readMM(paste0(path_dir_gDNA_reads,"/matrix.mtx.gz"))

dimnames(gDNA_mat_reads) <- list(gDNA_data_features_reads, gDNA_data_barcodes_reads)


#Calculate sum reads and sum features
gDNA_mat_reads_sum <- colSums(gDNA_mat_reads)
#gDNA_mat_reads_count <- apply(gDNA_mat_reads , 2, function(col) sum(col != 0))

gDNA_mat_reads_meta <- data.frame(nReads_gDNA = gDNA_mat_reads_sum, 
                                        #nFeature_gDNA_reads = gDNA_mat_reads_count, 
                                        cell = colnames(gDNA_mat_reads), 
                                        row.names = colnames(gDNA_mat_reads)) %>%
                                        replace(is.na(.), 0)



```



#Construct cDNA metadata table - from cDNA reads

```{r}

#load read data sets 

cDNA_data_features_reads <- read.delim(paste0(path_dir_cDNA_reads,"/features.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
cDNA_data_features_reads <- sub("_chrom", "\\1", cDNA_data_features_reads)
cDNA_data_barcodes_reads <- read.delim(paste0(path_dir_cDNA_reads, "/barcodes.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
cDNA_mat_reads <- readMM(paste0(path_dir_cDNA_reads, "/matrix.mtx.gz"))
dimnames(cDNA_mat_reads) <- list(cDNA_data_features_reads, cDNA_data_barcodes_reads)


#Make metatable for cDNA

#Calculate sum reads and sum features - include CROP-seq in total number of reads
cDNA_mat_reads_sum <- colSums(cDNA_mat_reads)


#Count features 
#cDNA_mat_reads_count <- apply(cDNA_mat_reads , 2, function(col) sum(col != 0))

cDNA_mat_reads_meta <- data.frame(nReads_RNA = cDNA_mat_reads_sum, 
                                        #nFeatures_RNA_reads = cDNA_mat_reads_count,
                                        cell_index = colnames(cDNA_mat_reads), 
                                        row.names = colnames(cDNA_mat_reads))


#Split sample BC info and assign
  
cell_id_split <- data.frame(do.call('rbind', strsplit(as.character(cDNA_mat_reads_meta$cell_index),':',fixed=TRUE)))
colnames(cell_id_split) <- c("cell", "sbc")
Types <- data.frame(Type = c("FixA_1", "FixA_2", "FixA_3", "FixA_4", "FixB_1", "FixB_2", "FixB_3", "FixB_4"),
                    sbc = c("TCGCCTTA", "CTAGTACG", "TTCTGCCT", "GCTCAGGA", "AGGAGTCC", "CATGCCTA", "GTAGAGAG", "CCTCTCTG"),
                    stringsAsFactors = FALSE)
cell_id_split <- left_join(cell_id_split , Types, by = "sbc")
cDNA_mat_reads_meta <- cbind(cDNA_mat_reads_meta, cell_id_split)


#Calculate max_sbc ratio - inlcude CROP reads here for ratio
cDNA_mat_reads_meta <- as.data.table(cDNA_mat_reads_meta)
max_sbc <- cDNA_mat_reads_meta[cDNA_mat_reads_meta[, .I[which.max(nReads_RNA)], by=cell]$V1]
sum_sbc <- cDNA_mat_reads_meta[, sum(nReads_RNA),by=list(cell)]
ratio_sbc <- max_sbc$nReads_RNA/sum_sbc$V1
max_sbc <- data.frame(cell_index = max_sbc$cell_index, cDNA_ratio_sbc = ratio_sbc, max_sbc = "max_sbc")

cDNA_mat_reads_meta <- cDNA_mat_reads_meta %>% left_join(max_sbc, by = "cell_index")


```


#Merge gDNA and cDNA metatable files 

```{r}


#Merge with gDNA file - ATTENTION - certain gDNA values duplicated as possibilyt that same cell has different sample barcodes - only keep max sbc for later - retain now to be compatible to make a seurat object

gDNA_cDNA_meta <- cDNA_mat_reads_meta %>% 
  left_join(gDNA_mat_reads_meta, by = c("cell")) %>%
                              replace(is.na(.), 0)


#Add a column for the sum of reads for both cDNA/CROP and gDNA - keep only max sbc later
gDNA_cDNA_meta$nReads_tot <- gDNA_cDNA_meta$nReads_RNA + gDNA_cDNA_meta$nReads_gDNA
row.names(gDNA_cDNA_meta) <- gDNA_cDNA_meta$cell_index


#Merge cDNA meta file base on reads only - disregard sbc info - merge with gNDA mat - no duplicates made this time 

cDNA_mat_reads_meta_wo_sbc <- cDNA_mat_reads_meta %>% 
                              group_by(cell) %>% 
                              summarise(nReads_RNA_wo_sbc = sum(nReads_RNA))%>%
                              left_join(gDNA_mat_reads_meta, by = c("cell"))  %>%
                              replace(is.na(.), 0)



#Calculate total number of reads
cDNA_mat_reads_meta_wo_sbc$nReads_tot <- cDNA_mat_reads_meta_wo_sbc$nReads_RNA_wo_sbc + cDNA_mat_reads_meta_wo_sbc$nReads_gDNA

cDNA_mat_reads_meta_wo_sbc <- cDNA_mat_reads_meta_wo_sbc %>% mutate(selection_status = if_else(nReads_tot >= 1000, "selected", "not_selected"))

#Make rank rank plot for cutoff threshold setting

group_colors <- c(selected = "#990000", not_selected = "black")

index_v_line <- cDNA_mat_reads_meta_wo_sbc %>%
  arrange(desc(nReads_tot)) %>%
  mutate(index = row_number()) %>%
  filter(nReads_tot == 1000) %>%
  summarize(last_index = last(index)) %>%
  pull(last_index)


title = "nReads_tot_ranked.png"
p <- cDNA_mat_reads_meta_wo_sbc %>% 
  arrange(desc(nReads_tot)) %>%
  mutate(index = 1:nrow(cDNA_mat_reads_meta_wo_sbc)) %>%
   ggplot() +
  	geom_point(aes(x=index, y= nReads_tot, color = selection_status)) +
    #geom_point(aes(x=index, y= nReads_gDNA), alpha = 0.2, color = "red")+
  #geom_point(aes(x=index, y= nReads_RNA_wo_sbc), alpha = 0.2, color = "darkgreen")+
  	theme_classic() +
    scale_x_log10() + 
  	scale_y_log10() + 
  geom_vline(xintercept = index_v_line, linetype = "longdash")+
      scale_fill_manual(values = group_colors)+
      scale_color_manual(values = group_colors)+
   guides(fill = "none", color = "none")+
  #xlim(c(0.3,0.7))+
  #	geom_vline(xintercept = 0.8)+
      theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)
     

    ggsave(filename = title, path = plot_save_dir, plot = p, device = "png", scale = 1,
       width = 15, height = 15, units = "cm", dpi = 300)



title = "nReads_RNA_ranked.png"
p <- cDNA_mat_reads_meta_wo_sbc %>% 
  arrange(desc(nReads_RNA_wo_sbc)) %>%
  mutate(index = 1:nrow(cDNA_mat_reads_meta_wo_sbc)) %>%
   ggplot() +
  	geom_point(aes(x=index, y= nReads_RNA_wo_sbc, color = selection_status)) +
    #geom_point(aes(x=index, y= nReads_gDNA), alpha = 0.2, color = "red")+
  #geom_point(aes(x=index, y= nReads_RNA_wo_sbc), alpha = 0.2, color = "darkgreen")+
  	theme_classic() +
    scale_x_log10() + 
  	scale_y_log10() + 
   # geom_vline(xintercept = index_v_line, linetype = "longdash")+
        scale_fill_manual(values = group_colors)+
      scale_color_manual(values = group_colors)+
   guides(fill = "none", color = "none")+
  #xlim(c(0.3,0.7))+
  #	geom_vline(xintercept = 0.8)+
      theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)


    ggsave(filename = title, path = plot_save_dir, plot = p, device = "png", scale = 1,
       width = 15, height = 15, units = "cm", dpi = 300)
  

    

title = "nReads_gDNA_ranked.png"
p <- cDNA_mat_reads_meta_wo_sbc %>% 
  arrange(desc(nReads_gDNA)) %>%
  mutate(index = 1:nrow(cDNA_mat_reads_meta_wo_sbc)) %>%
   ggplot() +
  	geom_point(aes(x=index, y= nReads_gDNA, color = selection_status)) +
    #geom_point(aes(x=index, y= nReads_gDNA), alpha = 0.2, color = "red")+
  #geom_point(aes(x=index, y= nReads_RNA_wo_sbc), alpha = 0.2, color = "darkgreen")+
  	theme_classic() +
    scale_x_log10() + 
  	scale_y_log10() + 
      #geom_vline(xintercept = index_v_line, linetype = "longdash")+
        scale_fill_manual(values = group_colors)+
      scale_color_manual(values = group_colors)+
   guides(fill = "none", color = "none")+
  #xlim(c(0.3,0.7))+
  #	geom_vline(xintercept = 0.8)+
      theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)

  
  ggsave(filename = title, path = plot_save_dir, plot = p, device = "png", scale = 1,
       width = 15, height = 15, units = "cm", dpi = 300)


```





## Make seurat from cDNA and add metadata


```{r}

#Laod matrix - umis
cDNA_data_features_umis <- read.delim(paste0(path_dir_cDNA_umis, "/features.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
cDNA_data_features_umis <- sub("_chrom", "\\1", cDNA_data_features_reads)
cDNA_data_barcodes_umis <- read.delim(paste0(path_dir_cDNA_umis, "/barcodes.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
cDNA_mat <- readMM(paste0(path_dir_cDNA_umis, "/matrix.mtx.gz"))
dimnames(cDNA_mat) <- list(cDNA_data_features_umis, cDNA_data_barcodes_umis)


#Make Seurat
cDNA_data_Seurat <- CreateSeuratObject(cDNA_mat, project = "SDR002_240_240")

#Add metadata
all.equal(colnames(cDNA_data_Seurat), rownames(gDNA_cDNA_meta))
cDNA_data_Seurat <- AddMetaData(object = cDNA_data_Seurat, metadata = gDNA_cDNA_meta)


# Only keep highest sbc - filter out contaminating sbcs
cDNA_data_Seurat <- subset(x = cDNA_data_Seurat, subset = max_sbc == "max_sbc")

metadata <- cDNA_data_Seurat@meta.data
#Rename cells to make merging with gDNA later easier 

all.equal(metadata$cell_index, colnames(cDNA_data_Seurat))


cDNA_mat_mod <- cDNA_data_Seurat@assays$RNA$counts
cells <- metadata$cell
features <- rownames(cDNA_data_Seurat)

dimnames(cDNA_mat_mod) <- list(features, cells)
rownames(metadata) <- cells

#Make Seurat
cDNA_data_Seurat <- CreateSeuratObject(cDNA_mat_mod, project = "SDR002_240_240")

#Add metadata
all.equal(colnames(cDNA_data_Seurat), rownames(metadata))
cDNA_data_Seurat <- AddMetaData(object = cDNA_data_Seurat, metadata = metadata)



```



#Cutoff to filter out low read cells

```{r}
# Filter out low quality cells based on reads - yield around 10000-12000 cells

#cDNA_data_Seurat <- subset(x = cDNA_data_Seurat, subset = nReads_tot >= 1000)

saveRDS(cDNA_data_Seurat, paste0(plot_save_dir,"/SDR002_240_240_Seurat.rds"))

metadata <- cDNA_data_Seurat@meta.data




```


###Add RNA reads as an assay 

```{r}

#load read data sets 

cDNA_data_features_reads <- read.delim(paste0(path_dir_cDNA_reads,"/features.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
cDNA_data_features_reads <- sub("_chrom", "\\1", cDNA_data_features_reads)
cDNA_data_barcodes_reads <- read.delim(paste0(path_dir_cDNA_reads, "/barcodes.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
cDNA_mat_reads <- readMM(paste0(path_dir_cDNA_reads, "/matrix.mtx.gz"))

#Only get cell barcode without sbc - only works here as cells are already filtered for max sbc - otherwise filtering for max sbc needs to be done
cell_id_split <- data.frame(do.call('rbind', strsplit(as.character(cDNA_data_barcodes_reads),':',fixed=TRUE)))

dimnames(cDNA_mat_reads) <- list(cDNA_data_features_reads, cell_id_split$X1)

#Get cells for subsetting
to_subset <- metadata$cell
valid_to_subset <- to_subset[to_subset %in% colnames(cDNA_mat_reads)]

# Check if all elements in to_subset are valid column names in cDNA_mat_reads
all(valid_to_subset %in% colnames(cDNA_mat_reads))

#Subset matrix and convert to df
cDNA_mat_reads_df <- cDNA_mat_reads[,valid_to_subset] %>% t() %>% as.data.frame()

#Add missing cells and convert to matrix again for adding 
#Make df of missing cells 
missing_cells <- setdiff(to_subset, valid_to_subset)


missing_cells_df <- matrix(0, nrow = length(missing_cells), ncol = length(colnames(cDNA_mat_reads_df)),
                           dimnames = list(missing_cells, colnames(cDNA_mat_reads_df)))
missing_cells_df <- as.data.frame(missing_cells_df)

#Combine the two dfs and brind in correct order 

cDNA_mat_reads_df <- rbind(cDNA_mat_reads_df, missing_cells_df)
cDNA_mat_reads_df <- cDNA_mat_reads_df[to_subset,]
all.equal(metadata$cell, rownames(cDNA_mat_reads_df))

cDNA_mat_reads_to_add <- cDNA_mat_reads_df %>% t()

#Add assay for gDNA
#Make assay
cDNA_assay <- CreateAssay5Object(counts = as.sparse(cDNA_mat_reads_to_add))
cDNA_data_Seurat[["RNA_reads"]] <- cDNA_assay


```



#Add gDNA as an assay to the Seurat

```{r}

#load read data sets 

gDNA_data_features_reads <- read.delim(paste0(path_dir_gDNA_reads, "/features.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
gDNA_data_features_reads <- sub("_chrom", "\\1", gDNA_data_features_reads)
gDNA_data_barcodes_reads <- read.delim(paste0(path_dir_gDNA_reads,"/barcodes.tsv.gz"), header = FALSE) %>% dplyr::pull(1)
gDNA_mat_reads <- readMM(paste0(path_dir_gDNA_reads,"/matrix.mtx.gz"))
dimnames(gDNA_mat_reads) <- list(gDNA_data_features_reads, gDNA_data_barcodes_reads)

#Get cells for subsetting
to_subset <- metadata$cell
valid_to_subset <- to_subset[to_subset %in% colnames(gDNA_mat_reads)]

# Check if all elements in to_subset are valid column names in gDNA_mat_reads
all(valid_to_subset %in% colnames(gDNA_mat_reads))

#Subset matrix and convert to df
gDNA_mat_reads_df <- gDNA_mat_reads[,valid_to_subset] %>% t() %>% as.data.frame()

#Add missing cells and convert to matrix again for adding 
#Make df of missing cells 
missing_cells <- setdiff(to_subset, valid_to_subset)


missing_cells_df <- matrix(0, nrow = length(missing_cells), ncol = length(colnames(gDNA_mat_reads_df)),
                           dimnames = list(missing_cells, colnames(gDNA_mat_reads_df)))
missing_cells_df <- as.data.frame(missing_cells_df)

#Combine the two dfs and brind in correct order 

gDNA_mat_reads_df <- rbind(gDNA_mat_reads_df, missing_cells_df)
gDNA_mat_reads_df <- gDNA_mat_reads_df[to_subset,]
all.equal(metadata$cell, rownames(gDNA_mat_reads_df))

gDNA_mat_reads_to_add <- gDNA_mat_reads_df %>% t()

#Add assay for gDNA
#Make assay
gDNA_assay <- CreateAssay5Object(counts = as.sparse(gDNA_mat_reads_to_add))
cDNA_data_Seurat[["gDNA"]] <- gDNA_assay

saveRDS(cDNA_data_Seurat, paste0(plot_save_dir,"/SDR002_240_240_Seurat.rds"))
cDNA_data_Seurat <- readRDS(paste0(plot_save_dir,"/SDR002_240_240_Seurat.rds"))

#Set plot order in metadata

#order <- c("PFA", "Glyoxal")
#metadata$Type <- factor(metadata$Type, levels = order)

#Set colors for plotting 

group_colors <- c(FixA_1 = "#EADFB4", FixA_2 = "#9BB0C1", FixA_3 = "#51829B" , FixA_4 = "#F6995C", 
                  FixB_1 = "#EFBC9B", FixB_2 = "#FBF3D5", FixB_3 = "#D6DAC8", FixB_4 = "#9CAFAA")

#group_colors <- c(PFA = "#003f5c", Glyoxal = "#ffa600")

#Calculate percentages for plotting 

metadata <- cDNA_data_Seurat@meta.data
metadata$nFeature_RNA_perc <- metadata$nFeature_RNA/panel_size
metadata$nFeature_gDNA_perc <- metadata$nFeature_gDNA/panel_size
avg_UMIs <- mean(metadata$nCount_RNA)

#cDNA_data_Seurat@meta.data <- metadata

```




###Visualize number of cells

```{r}

title = "NCells_per_Type.pdf"
p <- metadata %>%
  	ggplot(aes(x=Type, fill=Type)) + 
  	geom_bar() +
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    scale_y_continuous(limits = c(0,5000))+
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)+
    labs(x = "Type", y = "Number of cells")+
    scale_fill_manual(values = group_colors)+
    guides(fill = "none")


ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)
p

```

###Genes detected per cell - Violin

```{r}

title = "nGenes_Violin_perc.pdf"

p <- metadata %>% 
  	ggplot(aes(x=Type, y=nFeature_RNA_perc, fill=Type)) + 
  geom_violin(alpha = 0.2, adjust = 3)+
  geom_boxplot(width = 0.25)+
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
    scale_y_continuous(limits = c(0,1))+
  	ggtitle(title)+
      labs(x = "Type", y = "Genes per cell")+
    scale_fill_manual(values = group_colors)+
    guides(fill = "none")


ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)

p

```






###Genes detected per cell - Density 

```{r}

title = "nGenes_Density_perc.pdf"

p <- metadata %>% 
  	ggplot(aes(color=Type, x=nFeature_RNA_perc, fill= Type)) + 
  	geom_density(alpha = 0.2, adjust =3) + 
  	theme_classic() +
  	scale_x_continuous(limits = c(0, 1)) + 
  	geom_vline(xintercept = 10)+
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)+
      labs(x = "Type", y = "Genes per cell")+
    scale_fill_manual(values = group_colors)+
      scale_color_manual(values = group_colors)+
    guides(fill = "none", color = "none")

ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)


p
```



###Amplicons detected per cell - Violin

```{r}

title = "nAmplicons_Violin_perc.pdf"

p <- metadata %>% 
  	ggplot(aes(x=Type, y=nFeature_gDNA_perc, fill=Type)) + 
  geom_violin(alpha = 0.2, adjust = 3)+
  geom_boxplot(width = 0.25)+
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
    scale_y_continuous(limits = c(0,1))+
  	ggtitle(title)+
      labs(x = "Type", y = "Amplicons per cell")+
    scale_fill_manual(values = group_colors)+
    guides(fill = "none")


ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)

p

```



###Amplicons detected per cell - Density 

```{r}

title = "nAmplicons_Density_perc.pdf"

p <- metadata %>% 
  	ggplot(aes(color=Type, x=nFeature_gDNA_perc, fill= Type)) + 
  	geom_density(alpha = 0.2, adjust =3) + 
  	theme_classic() +
  	scale_x_continuous(limits = c(0, 1)) + 
  	geom_vline(xintercept = 10)+
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)+
      labs(x = "Type", y = "Genes per cell")+
    scale_fill_manual(values = group_colors)+
      scale_color_manual(values = group_colors)+
    guides(fill = "none", color = "none")

ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)


p
```


###UMIs detected per cell - Violin

```{r}


title = "nUMIs_Violin.pdf"

p <- metadata %>% 
  	ggplot(aes(x=Type, y=nCount_RNA, fill=Type)) + 
  geom_violin(alpha = 0.2, adjust = 1)+
  geom_boxplot(width = 0.25)+
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
    scale_y_continuous(limits = c(0,10000))+
  	ggtitle(title)+
      labs(x = "Type", y = "UMIs per cell (RNA)")+
    scale_fill_manual(values = group_colors)+
    guides(fill = "none")


ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)

p

```




###UMIs detected per cell - Density

```{r}

title = "nUMIs_Density_perc.pdf"

p <- metadata %>% 
  	ggplot(aes(color=Type, x=nCount_RNA, fill= Type)) + 
  	geom_density(alpha = 0.2, adjust =1) + 
  	theme_classic() +
  	scale_x_continuous(limits = c(0, 10000)) + 
  	geom_vline(xintercept = 0)+
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)+
      labs(x = "Type", y = "UMIs per cell (RNA)")+
    scale_fill_manual(values = group_colors)+
      scale_color_manual(values = group_colors)+
    guides(fill = "none", color = "none")

ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)


p



```


###Reads - Amplicon detected per cell - Violin

```{r}


title = "nReads_gDNA_Violin.pdf"

p <- metadata %>% 
  	ggplot(aes(x=Type, y=nReads_gDNA, fill=Type)) + 
  geom_violin(alpha = 0.2, adjust = 1)+
  geom_boxplot(width = 0.25)+
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
    scale_y_continuous(limits = c(0,60000))+
  	ggtitle(title)+
      labs(x = "Type", y = "Reads per cell (Amplicons)")+
    scale_fill_manual(values = group_colors)+
    guides(fill = "none")


ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)

p


```




###Reads - Amplicon detected per cell - Density

```{r}


title = "nAmplicons_Density_perc.pdf"

p <- metadata %>% 
  	ggplot(aes(color=Type, x=nReads_gDNA, fill= Type)) + 
  	geom_density(alpha = 0.2, adjust =1) + 
  	theme_classic() +
  	scale_x_continuous(limits = c(0, 10000)) + 
  	geom_vline(xintercept = 10)+
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)+
      labs(x = "Type", y = "Reads per cell (Amplicons)")+
    scale_fill_manual(values = group_colors)+
      scale_color_manual(values = group_colors)+
    guides(fill = "none", color = "none")

ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)


p



```


#Plot ratios of cell barcodes - sbc


```{r}
title = "cDNA_ratio_sbc.pdf"

#Plot sample barcode ratio
p <- metadata %>%
 ggplot(aes(x=Type, y=cDNA_ratio_sbc, fill = Type))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.25)+
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
    scale_y_continuous(limits = c(0.5,1))+
  	ggtitle(title)+
      labs(x = "Type", y = "% of maximum sample barcode per cell")+
    scale_fill_manual(values = group_colors)+
    guides(fill = "none")

p 

ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 10, height = 15, units = "cm", dpi = 300)



```


#Plot - gDNA

```{r}

title = "nCount_RNA-nReads_gDNA-nFeature_gDNA_perc_gDNA.pdf"
p <- metadata %>% 
  	ggplot(aes(x=nCount_RNA, y=nReads_gDNA, color=nFeature_gDNA_perc)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "#990000", limits = c(0,1)) +
  #xlim(c(0,30))+
    #ylim(c(0,30))+
  	scale_x_log10(limits = c(1,100000)) + 
  	scale_y_log10(limits = c(1,100000)) + 
  	theme_classic() +
  	#geom_vline(xintercept = 10) +
  	#geom_hline(yintercept = 10) +
    	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)
  	#facet_wrap(~Type)

p

ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 15, height = 15, units = "cm", dpi = 300)


```



#Plot - RNA

```{r}

title = "nCount_RNA-nReads_gDNA-nFeature_gDNA_perc_RNA.pdf"
p <- metadata %>% 
  	ggplot(aes(x=nCount_RNA, y=nReads_gDNA, color=nFeature_RNA_perc)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "#990000", limits = c(0,1)) +
  #xlim(c(0,30))+
    #ylim(c(0,30))+
  	scale_x_log10(limits = c(1,100000)) + 
  	scale_y_log10(limits = c(1,100000)) + 
  	theme_classic() +
  	#geom_vline(xintercept = 10) +
  	#geom_hline(yintercept = 10) +
    	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)
  	#facet_wrap(~Type)

p

ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 15, height = 15, units = "cm", dpi = 300)


```



#Plot - sbc

```{r}

title = "nCount_RNA-nReads_gDNA-nFeature_gDNA_perc_sbc.pdf"
p <- metadata %>% 
  	ggplot(aes(x=nCount_RNA, y=nReads_gDNA, color=cDNA_ratio_sbc)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "#990000", limits = c(0,1)) +
  #xlim(c(0,30))+
    #ylim(c(0,30))+
  	scale_x_log10(limits = c(1,100000)) + 
  	scale_y_log10(limits = c(1,100000)) + 
  	theme_classic() +
  	#geom_vline(xintercept = 10) +
  	#geom_hline(yintercept = 10) +
    	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle(title)
  	#facet_wrap(~Type)

p

ggsave(filename = title, path = plot_save_dir, plot = p, device = "pdf", scale = 1,
       width = 15, height = 15, units = "cm", dpi = 300)


```




